(Deep) Induction Rules for GADTs\vspace*{-0.1in}

### 2021-08-29 induction-paper/revised-induction-paper.tex
small tweaks

GADTs are used in precisely those situations in which different
behaviors at different instances of data types are desired. This is
achieved by allowing the programmer to give the type signatures of the
GADT's data constructors independently, and then taking advantage of
pattern matching to force the desired type refinement. For example,
the {\em equality} GADT
\begin{equation}\label{eq:equal}
\begin{array}{l}
\mathsf{data\ Equal : Set \to Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;\;\;\;\,refl :\, \forall \{A : Set\} \to Equal\,A\,A}
\end{array}
\end{equation}
is parameterized by two type indices, but it is only possible to
construct data elements of type $\mathsf{Equal\,A\,B}$ if $\mathsf{A}$
and $\mathsf{B}$ are instantiated at the same type. If the types
$\mathsf{A}$ and $\mathsf{B}$ are {\color{red} propositionally equal}
%syntactically identical
then the
type $\mathsf{Equal\,A\,B}$ contains the single data element
$\mathsf{refl}$. It contains no data elements otherwise.

@@

\noindent
This grammar is subject to the following restrictions. In the third
clause the type constructor $\mathsf{F_1}$ does not contain
$\mathsf{G}$. In the fourth clause, none of the $\mathsf{\alpha}$-many
type constructors in $\mathsf{\ol{F_1}}$ contains $\mathsf{G}$. This
prevents nesting, which would {\color{red} make it impossible to give}
an induction rule for $\mathsf{G}$; see Section~\ref{sec:GADT-nested}
below. In the fifth and sixth clauses, $\mathsf{H : Set^\gamma \to
  Set}$ is the syntactic reflection of some functor, and thus has an
associated map function. It is worth noting that the fifth clause
subsumes the cases in which $\mathsf{F\,G\,\ol{B}}$ is a closed type
or one of the $\mathsf{B_i}$, and that $\mathsf{H}$ can be the data
type constructor for any (truly) nested type. From the map function
for $\mathsf{H}$ we can also construct a map function

>>>>

GADTs are used in precisely those situations in which different
behaviors at different instances of data types are desired. This is
achieved by allowing the programmer to give the type signatures of the
GADT's data constructors independently, and then taking advantage of
pattern matching to force the desired type refinement. For example,
the {\em equality} GADT
\begin{equation}\label{eq:equal}
\begin{array}{l}
\mathsf{data\ Equal : Set \to Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;\;\;\;\,refl :\, \forall \{A : Set\} \to Equal\,A\,A}
\end{array}
\end{equation}
is parameterized by two type indices, but it is only possible to
construct data elements of type $\mathsf{Equal\,A\,B}$ if $\mathsf{A}$
and $\mathsf{B}$ are instantiated at the same type. If the types
$\mathsf{A}$ and $\mathsf{B}$ are syntactically identical then the
type $\mathsf{Equal\,A\,B}$ contains the single data element
$\mathsf{refl}$. It contains no data elements otherwise.

@@

\noindent
This grammar is subject to the following restrictions. In the third
clause the type constructor $\mathsf{F_1}$ does not contain
$\mathsf{G}$. In the fourth clause, none of the $\mathsf{\alpha}$-many
type constructors in $\mathsf{\ol{F_1}}$ contains $\mathsf{G}$. This
prevents nesting, which would make it impossible to give an induction
rule for $\mathsf{G}$; see Section~\ref{sec:GADT-nested} below. In the
fifth and sixth clauses, $\mathsf{H : Set^\gamma \to Set}$ is the
syntactic reflection of some functor, and thus has an associated map
function. It is worth noting that the fifth clause subsumes the cases
in which $\mathsf{F\,G\,\ol{B}}$ is a closed type or one of the
$\mathsf{B_i}$, and that $\mathsf{H}$ can be the data type constructor
for any (truly) nested type. From the map function for $\mathsf{H}$ we
can also construct a map function
